### 2011-11-04 doc/pldi2012/highlevel.tex

\subsubsection{Dependence Detection}
\label{subsec:depdetect}
When a task is registered, the runtime
will perform dependence detection for every region that the task requests with
prior tasks that have been registered in the same parent task context.  Since
tasks are registered in program order we will always perform dependence detection
in program order which guarantees correctness even if tasks are executed out-of-order.  
Performing dependence detection in order is also annalogous to how an out of order
processor works.  Out-of-order processors load instructions in order and only after
analyzing their instructions and register dependencies do they execute instructions
out of order.  The regions that a task requires are annalogous
to the registers listed in an instruction and the goal of the high level
runtime is to detect the same dependences and use them to enforce scheudling constraints
upon a task.  However, there is a catch to this analogy: in the Legion programming
model two different logical regions are not necessarily independent from each other.  In
order to do the dependence detection we rely on the region tree data structure to be
able to detect dependences between tasks which require different logical regions.

@@

Once we have found the root region for the requested region, we then proceed to walk
down the region tree from the node corresponding to the root region to the node
for the requested region.  At each node, we determine if any potential dependences exist.

For region nodes we first check to see if the required region has any dependences on
tasks currently using the logical region by performing the check on coherence properties
described in section \ref{subsec:cohdep}.  If a conflict is discovered
we record that the task must wait on the previous task before executing.  We also check
to see if any of the partitions of the region have been marked open.  If they have
we must {\em close} the partitions to detect any possible conflicts

\subsection{Mapping}

>>>>

\subsubsection{Dependence Detection}
\label{subsec:depdetect}
When a task is registered, the runtime
will perform dependence detection for every region that the task requests with
prior tasks that have been registered in the same parent task context.  Since
tasks are registered in program order we will always perform dependence detection
in program order which guarantees correctness even if tasks are executed out-of-order.  
Performing dependence detection in program order is also annalogous to how an out of order
processor works.  Out-of-order processors load instructions in order and only after
analyzing their instructions and register dependencies do they execute instructions
out of order.  The regions that a task requires are annalogous
to the registers listed in an instruction and the goal of the high level
runtime is to detect the same dependences and use them to enforce scheudling constraints
upon a task.  However, there is a catch to this analogy: in the Legion programming
model two different logical regions are not necessarily independent from each other.  In
order to do the dependence detection we rely on the forest of region tree data structures to be
able to detect dependences between tasks which require different logical regions.

@@

Once we have found the root region for the requested region, we then proceed to walk
down the region tree from the node corresponding to the root region to the node
for the requested region.  At each node, we determine if any potential dependences exist.
Determining whether dependencies exist involves checking information both about the current
node being visited as well as whether there are any potential dependences farther down
in the region tree.  We now describe the decision procedure performed at both region
nodes and partition nodes in the region trees.

For region nodes, we begin by checking if there are any tasks actively using the logical
region which we depend.  We determine this using the coherence dependence analysis described 
in section \ref{cohdep}.  For each dependence that is found, we record that the previous
task must have completed prior to the start of the task we are registering.  After performing
dependence analysis we then check if there is an open partition which may contain additional
dependencies.  If the open partition is the partition we intend to traverse then we
simply continue the traversal.  However, if the partition we want is not open then we must
first {\em close} up any open partition and then open the partition we are going to
traverse.  The act of closing a partition involves traversing the region tree and recording
all active tasks as dependences.  The reason for this is that we have no disjointness
information about tasks that use different partitions and therefore to be safe must register
all task touching subregions as dependences.  After closing any open partitions, we
can then open the partition on our path and continue the traversal.

When we reach our target logical region we again perform the coherence dependence analysis.
We also close up any open partitions as they might contain tasks which could contain
a dependence.  If our task was dependent on all of the previous active tasks in a region
we say that our task was a {\em dominator} of all the previous active tasks.  We can
then clear the list of active tasks and replace it with the current task.  This is safe since
any later tasks that depend on the current task will have a transitive dependence 
through the dominator task.  If our task is not a dominator task, then we simply append
it to the list of active tasks making it possible for future task registrations to detect
it as a dependence.

When traversing partition nodes we perform a similar, but less complicated analysis.  There
are two cases to consider for partition nodes: disjoint partitions and aliased partitions.
In the case of disjoint partitions, we simply open the subregion that we intend to traverse
and then traverse it.  This is correct because we know by definition that any two logical
subregions in a disjoint partition are disjoint and therefore a dependence can never exits
between them.

For aliased partitions we have to perform an additional dependence analysis.  At every
aliased partition node we maintain a list of tasks active in all of the subregions.  We
perform a coherence dependence analysis between every pair of tasks regardless of the subregion
they are accessing to see if a dependence could exist.  We do this regardless of the subregion
because we have no disjointness information in an aliased partition.  If there are any
potential dependences, we close up all the subregions and then open the subregion required.  If
no dependence was detected then we open the subregion we need, append the task to the list
of active tasks, and continue the traversal.  Allowing multiple open subregions in an aliased
partition is safe because we performed an analysis between all active tasks regardless of
whether the particular subregions were aliased or not.

\subsubsection{Putting It All Together: Circuit Example}
\label{subsec:cirdependence}

\subsection{Implicit Dataflow Construction}
\label{subsec:dataflow}
In addition to computing the dependences between tasks, we also have to compute where
a task is permitted to acquire its data from when it goes to execute.
It is insufficient to simply record the producer task as there
may be multiple active tasks for a logical region, all of which may produce different physical
instances.  In addition, it is likely that the producer tasks have not been mapped
yet which makes it impossible to know the set of physical instances from which to pull data.

To solve this problem we create the notion of an {\em abstract instance}.  An abstract instance
is a place holder for the set of physical instances from which a task can choose to pull
data when it creates its own physical instances before executing.  For each region in
a task we record the source and destination abstract instances that correspond to the physical
instances from which a task will read its input data and then place its output data.
When the task is mapped as described in section \ref{subsec:mapping} we will update these
abstract instances with the actual physical instances chosen.

To determine the source and destination abstract instances of a region, we keep track of 
the previous abstract instances that were observed on the way to our target region node
when performing dependence analysis.  These abstract instances describe the set of physical
instances from which valid data for our task can be chosen when it is mapped.  If a region
subtree must be closed then all the abstract instances in that subtree are marked closed
so no additional tasks can register them as sources.  We can also chose to create a new
abstract instance to replace a currently valid abstract instance to avoid a WAR conflict.
By creating an abstract instance we will force the mapper to choose a different physical
instance from the existing ones when mapping is performed.

Abstract instances correspond to a dynamic construction of the dataflow graph.  Each abstract
instance is a node in the dataflow graph corresponding to a specific version of a logical 
region and its instantiations as physical regions.  Tasks are then edges between abstract
instances indicating operations on different versions of data.  By computing this information
during conflict detection, we can record it before tasks begin executing out of order allowing
us to guarantee program correctness in the presence of out of order task execution.  Continuing
the analogy with out-of-order processors, this implicit construction of the dataflow graph
is annalogous to register renaming which is done in order, but allows the processor to remove
WAR dependences and execute instructions in parallel that may have had false dependences.

\subsection{Mapping}
\label{subsec:mapping}
